---
title: "P1_RasterIntroduction"
author: "João Gonçalves"
date: "17 November 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.path = "img/")
```

Geospatial data is becoming increasingly used to solve numerous 'real-life' problems such as:
( ... examples here ...). In turn, R is becoming more equipped than ever to handle this type of data 
thus providing an exceptional open-source solution to solve many problems in the Geographic Information 
Sciences and Remote Sensing domains. 

In general two types of geospatial data models are used to represent, visualize and model spatial 
phenomena, these are: 

- __Vector data__: which represents the world in three simple geometries: __points__, __lines__ 
and __polygons__. As such, it allows to represent spatial phenomena or variables that are typically 
discrete and with  well-defined boundaries (e.g., touristic points-of-interest, gas stations, 
rivers, roads, drainage basins, country GDP).

- __Raster data__: provides support for representing spatial phenomena by diving the surface 
into a grid (or matrix) composed by cells of regular size. Each raster dataset has a certain 
number of columns and rows and each cell contains a value with information for the variable of 
interest. Stored data can be either: (i) thematic - representing a _discrete_ variable (e.g., 
land use/cover) or _continuous_ (e.g., elevation).

Choosing the appropriate data model to use depends on the domain of application and the specific 
problem at hand. Typically, people from the social sciences tend to use more the vector data model. 
R packages such as __sp__ or __sf__ (a relatively new package, starting in 2016), provide support for 
this type of data. In contrast, in the environmental sciences the raster data model is more often 
used because of satellite data or the need to represent spatially continuous phenomena such as 
pollution levels, temperature or precipitation values, the abundance or habitat suitability for 
a species among many other. The __raster__ package, introduced in March 2010 by Robert Hijmans 
& Jacob van Etten, currently provides many useful functions for using this type of data. Despite 
these differences, GIS specialists and researchers often use both data models to tackle their 
problems.

Throughout these posts we will cover the basics, intermediate and some advanced stuff in __raster data__ 
handling, manipulation and modelling in R. Examples will be given as well as exercises with these different 
levels will be presented so you can practice.



```{r install_packages, message=FALSE, warning=FALSE}


if(!("raster" %in% installed.packages()[,1])) install.packages(raster)

library(raster)

hist(rnorm(1000))


```

<center>![image-1](http://r-exercises.com/wp-content/uploads/2016/11/)</center>



download.file("https://raw.githubusercontent.com/joaofgoncalves/R_exercises_raster_tutorial/master/data/srtm_pnpg.tif",
              "srtm.tif", method="auto")







